---
title: "WallMart Coursework in R"
output: html_notebook
---


Installing Packages and Libraries
```{r}
install.packages("data.table")
library(data.table)
install.packages("ggplot2")
library(ggplot2)
install.packages("psych")
library(psych)
install.packages("corrplot")
library(corrplot)
install.packages("dplyr")
library(dplyr)
install.packages("plyr")
library(plyr)
install.packages("Amelia")
library(Amelia)
library(tidyr)
library(stringr)
library(dummies)
```

Setting the Working Directory & Loading the datasets as data frames
```{r}
#setwd("D:/RCW/")
train <- read.csv("D:/RCW/Train.csv")
test = read.csv("D:/RCW/Test.csv")
```

Printing the top 6 rows of the train & test data frames by using head()
```{r}
head(train)
```

```{r}
head(test)
```


Checking the Dimensions of the dataset by using dim()
The train dataset has 8523(rows) 12(cols)
The test dataset has 5681(rows) 11(cols)
```{r}
dim(train)
dim(test)
```

Checking the column names to find the missing columns
We find the 'Item_Outlet_Sales' is missing from test dataset.
This is because we will be predicting the values of 'Item_Outlet_Sales'
```{r}
names(train)
```
```{r}
names(test)
```

Checking if this data has missing values.
We are using table to group the values by False and True.
We find 1463 NA values in the train dataset.
```{r}
table(is.na(train))
```
Checking the variables wisth the count of NA values.
We find only Item_Weight has the 1463 NA values.
```{r}
colSums(is.na(train))
```

Checking the variables and their types in train dataset
```{r}
str(train)
```

We will see a summary of the train dataset.
```{r}
summary(train)
```

From the above information we can:

Exploring the Numerical Columns:

1.Item_Weight - There are 1463 NA Values

2.Item Visibility - Contains no NA Values, but contains 0 values.

3.Item_MRP - Contains No NA/0 values.Also has an acceptable price range with no outliers.

4.Outlet_Establishment_Year - Contains no NA/0 values. Average mean is 1997, implying mostly old stores.

5.Item_Outlet_Sales - Contains no NA/0 values.


Exploring the Factor Columns:

1.Item_Identifier - Contains 1559 unique values
```{r}
install.packages("plyr")
library(plyr)
library(dplyr)
train %>% 
  summarise(n_distinct(Item_Identifier))
```

2.Item_Fat_Content - We find the level values Low Fat/low fat/LF are same but typed incorrectly. 
```{r}
train %>%
group_by(Item_Fat_Content) %>% summarise(Count = n()) %>% arrange(desc(Count))
```

3.Item_Type - Categories of Items with counts
```{r}
train%>%
group_by(Item_Type) %>%
summarise(Count = n()) %>% arrange(desc(Count))
```

4.Outlet_Identifier - There are Item information from 10 different Outlets

```{r}
train %>%
group_by(Outlet_Identifier) %>%
summarise(Count = n()) %>% arrange(desc(Count))
```


5.Outlet_Size - Outlet Size data not properly levelled. (2410 counts)

```{r}
train%>%
group_by(Outlet_Size) %>% summarise(Count = n())
```

6.Outlet_Location_Type - Number of Outlet Location type with counts. We find the data is normally distributed.

```{r}
train%>%
group_by(Outlet_Location_Type) %>%
summarise(Count = n()) %>% arrange(desc(Count))
```


7.Outlet_Type - We find the Types of Outlet
```{r}
train%>%
group_by(Outlet_Type)%>%
summarise(Count=n())%>% arrange(desc(Count))
```



Data Manipulation


We are creating a new variable in test dataset Item_Outlet_Sales, to match our number of rows with train dataset.
```{r}
test$Item_Outlet_Sales <- 1
names(test)
```

Now, we are combining thee train and test data with rbind function
```{r}
combi <- rbind(train, test)
dim(combi)
```


We are imputing the NA values in Item_Weight with the median of the values of the column. 
To calculate the median of the non-missing values if are passing the argument na.rm=TRUE 
```{r}

combi$Item_Weight[is.na(combi$Item_Weight)] <- median(combi$Item_Weight, na.rm = TRUE)
summary(combi$Item_Weight)

```

There are 0 values in in Item_visibility, therefore, we also impute the 0s with median of the column values.
```{r}
combi$Item_Visibility <- ifelse(combi$Item_Visibility == 0, median(combi$Item_Visibility),combi$Item_Visibility)
summary(combi$Item_Visibility)

```

Renaming the blank level in of Outlet_Size to 'Other'
```{r}
levels(combi$Outlet_Size)[1] <- "Other"
table(combi$Outlet_Size)
```

Renaming the levels of Item_Fat_Content to 'Low Fat' & 'Regular'
```{r}
combi$Item_Fat_Content <- revalue(combi$Item_Fat_Content,c("LF" = "Low Fat", "reg" = "Regular", "low fat" = "Low Fat"))
table(combi$Item_Fat_Content)

```



Data Visualisation


```{r}
str(combi)
```
Dividing data to Train & Test before Label & Hot Encoding

```{r}
new_train_combi <- combi %>% filter(Item_Outlet_Sales != 1)
new_test_combi <- combi %>% filter(Item_Outlet_Sales == 1)
str(new_train_combi)
str(new_test_combi)
```


We have tried to visualise Item_Outlet_Sales with different Categorical Values

```{r}

ggplot(new_train_combi, aes(Item_Fat_Content, Item_Outlet_Sales)) + geom_bar(stat = "identity", color = "red") + ggtitle("Item_Fat_Content vs Total Sales")

ggplot(new_train_combi, aes(Item_Type, Item_Outlet_Sales)) + geom_bar(stat = "identity", color = "red") + theme(axis.text.x = element_text(angle = 90), axis.text.y = element_text(angle = 0)) + ggtitle("Item_Type vs Total Sales")

ggplot(new_train_combi, aes(Outlet_Size, Item_Outlet_Sales)) + geom_bar(stat = "identity", color = "blue") + ggtitle("Outlet_Size vs Total Sales")

ggplot(new_train_combi, aes(Outlet_Location_Type, Item_Outlet_Sales)) + geom_bar(stat = "identity", color = "red") + ggtitle("Outlet_Location_Type vs Total Sales")

ggplot(new_train_combi, aes(Outlet_Type, Item_Outlet_Sales)) + geom_bar(stat = "identity", color = "blue") + ggtitle("Outlet_Size vs Total Sales")

qplot(x=Item_Type,y=Item_Outlet_Sales, data=new_train_combi,geom = "boxplot",) + theme(axis.text.x = element_text(angle = 90), axis.text.y = element_text(angle = 0))

ggplot(combi, aes(Item_Type, Item_MRP)) + geom_bar(stat = "identity", color = "blue")+ theme(axis.text.x = element_text(angle = 90), axis.text.y = element_text(angle = 0)) + ggtitle("Item_Type vs Item_MRP")

```



We have tried to visualise Item_Outlet_Sales with different Continuous Values:


```{r}
ggplot(new_train_combi, aes(Item_Weight, Item_Outlet_Sales)) + geom_point(size = .5, color="navy") + ggtitle("Item_Weight vs Item Outlet Sales")

ggplot(new_train_combi, aes(Item_Visibility, Item_Outlet_Sales)) + geom_point(size = .5,  color = "blue") + ggtitle("Item_Visibility vs Total Sales")

ggplot(new_train_combi, aes(Outlet_Establishment_Year, Item_Outlet_Sales)) + geom_bar(stat = "identity", color="blue")  + ggtitle("Outlet_Establishment_Year vs Item Outlet Sales")

```




Manupulating data with Label Encoding & Hot Encoding.

Creating a dataframe combi_encoded similar to combi. 
```{r}
combi_encoded=as.data.frame(combi)
str(combi_encoded)
```


Label Encoding.
We will change the categorical variable Item_Fat_Content to numeric 0 & 1.
```{r}
combi_encoded$Item_Fat_Content <- ifelse(combi_encoded$Item_Fat_Content == "Regular",1,0)
str(combi_encoded)
```


Hot Encoding.
We will use dummy.data.frame() to split the catrgorical variable to a matrix of variables 0 and 1, 
```{r}
library(dummies)
combi_encoded_dummies <- dummy.data.frame(combi_encoded, names = c('Outlet_Size','Outlet_Location_Type','Outlet_Type'),sep = '_')

str(combi_encoded_dummies)
```


Now, We will save the dataframe with all columns with int and num and Drop thecolumns with Categorical variables/Factors.

Item_Identifier, Outlet_Identifier,Item_Type has not been converted to matrix because of the high no of factor levels, which we cannot compute due to limited system resources.

Hence we are removing the 3 variables from the final variable.
```{r}
combi_encoded_dummies_drop <- select(combi_encoded_dummies, -c(Item_Identifier, Outlet_Identifier,Item_Type))
str(combi_encoded_dummies_drop)
```

```{r}
summary(combi_encoded_dummies_drop)
```

Dividing data to Train & Test POST Label & Hot Encoding
```{r}
new_train_combi_encoded_dummies_drop <- combi_encoded_dummies_drop %>% filter(Item_Outlet_Sales != 1)
new_test_combi_encoded_dummies_drop <- combi_encoded_dummies_drop %>% filter(Item_Outlet_Sales == 1)
str(new_train_combi_encoded_dummies_drop)
str(new_test_combi_encoded_dummies_drop)
```



```{r}
library(GGally)
ggpairs(combi_encoded_dummies_drop)
```

















